---
title: "class11"
author: "Katie Chau"
date: "2/28/2022"
output:
  pdf_document: default
  html_document: default
---

```{r}
#install.packages("BiocManager")

```

```{r}
#BiocManager::install()

```
```{r}
#BiocManager::install("DESeq2")
```
```{r}
library(BiocManager)
library(DESeq2)
```
## Import countData and colData 

```{r}
counts <- read.csv("airway_scaledcounts.csv", row.names=1)
metadata <-  read.csv("airway_metadata.csv")
```


```{r}
head(counts)
```

```{r}
head(metadata)
```
```{r}
View(counts)
```
```{r}
View(metadata)
```

>Q1. How many genes are in this dataset? 

38694

>Q2. How many 'control' cell lines do we have? 

```{r}
dim(counts)
```
```{r}
dim(metadata)
```
4 control cell lines

## Toy Differential Gene Expression 
```{r}
control <- metadata[metadata[,"dex"]=="control",]
control.counts <- counts[ ,control$id]
control.mean <- rowSums( control.counts )/4 
head(control.mean)
```

I prefer this method over dyplr because I'd have to install the dyplr package, whereas this method is something we learned earlier so I'm more familiar with it. 

>Q3. How would you make the above code in either approach more robust?

You can put the counts[,control$id] into the rowSums parenthesis instead of putting it into control.counts. 

>Q4. Follow the same procedure for the treated samples (i.e. calculate the mean per gene across drug treated samples and assign to a labeled vector called treated.mean)

```{r}
treated <- metadata[metadata[,"dex"]=="treated",]
treated.counts <- counts[ ,treated$id]
treated.mean <- rowSums( treated.counts )/4 
head(treated.mean)
names(treated.mean) <- counts$ensgene
```

```{r}
meancounts <- data.frame(control.mean, treated.mean)
colSums(meancounts)
```

>Q5. (a). Create a scatter plot showing the mean of the treated samples against the mean of the control samples. Your plot should look something like the following.

```{r}
plot(meancounts, xlab= "Control", ylab = "Treated")
```


>Q5 (b).You could also use the ggplot2 package to make this figure producing the plot below. What geom_?() function would you use for this plot?

You would use geom_point() with the ggplot2 package. 

>Q6. Try plotting both axes on a log scale. What is the argument to plot() that allows you to do this?

```{r}
plot(meancounts, log= "xy")

```
log = "xy" allows you to plot the log of both axis with plot(). 

```{r}
#calculate log2foldchange and add to meancounts
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
```
>Q7. What is the purpose of the arr.ind argument in the which() function call above? Why would we then take the first column of the output and need to call the unique() function?

The arr.ind argument returns only TRUE values and in this case, it will only return non zeroes of the genes. 


>Q8. Using the up.ind vector above can you determine how many up regulated genes we have at the greater than 2 fc level?

250 

```{r}
up.ind <- mycounts$log2fc > 2
down.ind <- mycounts$log2fc < (-2)
```

```{r}
sum(up.ind, na.rm= TRUE)
```


>Q9. Using the down.ind vector above can you determine how many down regulated genes we have at the greater than 2 fc level?


```{r}
sum(down.ind, na.rm=TRUE)
```

367 

>Q10. Do you trust these results? Why or why not?

Yes because it calls back all the TRUE greater than 2 fc level down regulated genes. 

## DESeq2 Analysis 


```{r}
library(DESeq2)
citation("DESeq2")
```

# Importing Data 
```{r}
# Design formula 
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex)
dds
```

# DEseq Analysis 

```{r}
dds <- DESeq(dds)
```
# Getting Results 
```{r}
res <- results(dds)
res
```
```{r}
#Convert res to dataframe 
resDf <- as.data.frame(res)
View(resDf)
```

When the average overall expression of the gene is 0, the p-value is NA. 


```{r}
summary(res)
```


```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```


## Adding Annotation Data 

```{r}
library("AnnotationDbi")
```
```{r}
#BiocManager::install("org.Hs.eg.db")
```


```{r}
library("org.Hs.eg.db")
```
```{r}
columns(org.Hs.eg.db)
```

```{r}
res$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(res),     # Our gene names
                     keytype="ENSEMBL",        # The format of our gene names
                     column="SYMBOL",          # The new format we want to add
                     multiVals="first")
```
```{r}
head(res)
```

>Q11. Run the mapIds() function two more times to add the Entrez ID and UniProt accession and GENENAME as new columns called res$entrez, res$uniprot and res$genename.

```{r}
res$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     keytype="ENSEMBL", 
                     column="ENTREZID",          
                     multiVals="first")
```

```{r}
res$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(res),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")
```

```{r}
res$genename <- mapIds(org.Hs.eg.db,
                     keys=row.names(res), 
                     keytype="ENSEMBL", 
                     column="GENENAME",          
                     multiVals="first")
```

```{r}
head(res)
```


```{r}
# Adjusted P-Value
ord <- order( res$padj )
#View(res[ord,])
head(res[ord,])
```

```{r}
write.csv(res[ord,], "deseq_results.csv")
```

## Data Visualization 

#Volcano Plot 

```{r}
plot( res$log2FoldChange,  -log(res$padj), 
      xlab="Log2(FoldChange)",
      ylab="-Log(P-value)")
```

```{r}
plot( res$log2FoldChange,  -log(res$padj), 
 ylab="-Log(P-value)", xlab="Log2(FoldChange)")

# Add some cut-off lines
abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.05), col="darkgray", lty=2)
```

```{r}
# setup our custom point color vector 
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ]  <- "red" 

inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

# volcano plot with custom colors 
plot( res$log2FoldChange,  -log(res$padj), 
 col=mycols, ylab="-Log(P-value)", xlab="Log2(FoldChange)" )

# cut-off lines
abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)
```
Enhanced Volcano 

```{r}
#BiocManager::install("EnhancedVolcano")
```


```{r}
library(EnhancedVolcano)
```

```{r}
x <- as.data.frame(res)
EnhancedVolcano(x,
    lab = x$symbol,
    x = 'log2FoldChange',
    y = 'pvalue')
```
## Pathway Analysis 

Used to reduce complexity of gen list interpretation. 

# Pathway Analysis with R and Bioconductor 

Use GAGE package to do KEGG pathway enrichment 

```{r}
library(pathview)
library(gage)
library(gageData)

data(kegg.sets.hs)

# Examine the first 2 pathways in this kegg set for humans
head(kegg.sets.hs, 2)
```

```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```

```{r}
keggres = gage(foldchanges, gsets=kegg.sets.hs)
```

```{r}
attributes(keggres)
```

```{r}
# Look at the first three down (less) pathways
head(keggres$less, 3)
```
```{r}
pathview(gene.data=foldchanges, pathway.id="hsa05310")
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa05310", kegg.native=FALSE)
```

>Q12. Can you do the same procedure as above to plot the pathview figures for the top 2 down-reguled pathways?

yes
