---
title: "Covid-19 Vaccination Rates Mini-Project"
author: "Katie Chau"
date: "3/5/2022"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document: default
---

```{r}
# import vaccination data 
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```
>Q1. What column details the total number of people fully vaccinated? 

persons_fully_vaccinated 


>Q2. What column details the Zip code tabulation area?

zip_code_tabulation_area

>Q3. What is the earliest date in this dataset?

2021-01-05

>Q4. What is the latest date in this dataset? 

```{r}
nrow(vax)
```

The number of rows in the mini-project site was 105840, so to get the right answer on the website, I used that row's date and not the date that I got when I downloaded the csv file with more information. 
```{r}
vax$as_of_date[105840]
```

```{r}
vax$as_of_date[107604]
```
2022-03-01

```{r}
#Overview of dataset 
#install.packages("skimr")
skimr::skim(vax)
```
>Q5. How many numeric columns are in this dataset?

10 (on the website it has 9 as the correct answer but for my data as well as the data retrieved from skim(vax) on the website it says 10)


>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 


18338 (18174 on the website)

```{r}
sum( is.na(vax$persons_fully_vaccinated) )
```


> What percent of persons_fully_vaccinated values are missing (to 2 significant figures)?

```{r}
18174/77213*100
```

I tried this calculation for the data on the website but I couldn't get the correct number to show green. 

```{r}
#The same equation but with my data set.
18338/77744*100
```

>Q8. [Optional]: Why might this data be missing?

This data may be missing because this many people in the population aren't vaccinated. This could be due to being too young to receive a vaccination since a person must be 5 years or older. 


#Working with dates
```{r}
#install.packages("lubridate")
```

```{r}
library(lubridate)
```

```{r}
today()
```

```{r}
#error
#today() - vax$as_of_date[1]
```


```{r}
#specify using year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
#how many days since first vaccination reported in dataset
today() - vax$as_of_date[1]
```

```{r}
#days in dataset span
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

> Q9. How many days have passed since the last update of the dataset?

```{r}
vax$as_of_date[nrow(vax)] - today()
```

6 days have passed since the last update of the dataset.

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 

```{r}
library(dplyr)
n_distinct(vax$as_of_date)
```

61 unique dates are detailed in my data set. For the website, the answer is 60 for that dataset.

# Working with ZIP codes

```{r}
#library(zipcodeR)
```

```{r}
#install.packages("sf")
```
```{r}
#install.packages("gdal-config")
```

Cannot load zipcodeR package.Kept getting error " there is no package called 'sf'. Couldn't install package 'sf' due to gdal-config not found. gdal-config would not install for this version of R. 

```{r}
#distance between centroids of any two zip codes in miles
#geocode_zip('92037')
#zip_distance('92037','92109')
```
```{r}
#census data on zip code areas
#reverse_zipcode(c('92037', "92109") )
```
```{r}
# Pull data for all ZIP codes in the dataset
#zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```

## Focus on the San Diego Area 

```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```

```{r}
sd.10 <- filter(vax, county == "San Diego" &
                age5_plus_population > 10000)
```




```{r}
distinctZip <- filter(vax, county == "San Diego")
uniqueZip <- unique((distinctZip))
nrow(uniqueZip)
```


107 distinct zip codes in San Diego. Not sure why my output does not show the correct info with the unique function. 

```{r}
which.max(distinctZip$age12_plus_population)
```


```{r}
distinctZip[91,]
```
92154 is the zip code with highest age12 plus population. 


> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-02-22”?

```{r}
pfv <- filter(vax, county == "San Diego", as_of_date == "2022-02-22")
pfv
```


```{r}
noNA <- pfv$percent_of_population_fully_vaccinated[!is.na(pfv$percent_of_population_fully_vaccinated)]
mean(noNA)
```

0.7041551 is the average "Percent of Population Fully Vaccinated" in San Diego county as of 2022-02-22. 

```{r}
library(ggplot2)
```
```{r}

ggplot(pfv, aes(pfv$percent_of_population_fully_vaccinated,)) + geom_bar()
```

# Focus on UCSD/La Jolla
```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```


```{r}
ggplot(ucsd) +
  aes(x=as_of_date,
   y= percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")
```
## Comparing to similar sized areas 

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-02-22")

#head(vax.36)
```

> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?


```{r}
vaxNew <- vax.36$percent_of_population_fully_vaccinated[!is.na(vax.36$percent_of_population_fully_vaccinated)]
hline <- mean(vaxNew)
hline
```
```{r}
ggplot(ucsd) +
  aes(x=as_of_date,
   y= percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent Vaccinated")+
  geom_hline(yintercept = hline)
```
> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-02-22”?

```{r}
summary(vaxNew)
```

> Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36)+
  aes(x=percent_of_population_fully_vaccinated)+
  geom_bar()
```
> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2022-02-22") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)
```
```{r}
vax %>% filter(as_of_date == "2022-02-22") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
Both are below the average value for the calculated mean of the 92037 zip code that we calculated above. 

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color= "red") +
  ylim(0,1) +
  labs(x="Date", y="Percent Vaccinated",
       title="Vaccination Rates Across California",
       subtitle="Only areas with population above 36k are shown") +
  geom_hline(yintercept = hline, linetype="dashed")

```



> Q21. How do you feel about traveling for Spring Break and meeting for in-person class afterwards?

I would like things to go back to normal but I am scared with going back in person. 
