---
title: "Machine Learning 1"
author: "Katie Chau"
date: "2/14/2022"
output: pdf_document
---

# First up kmeans()

Demo of using kmeans() function in base R. First make up some data with a known structure.  
```{r}
tmp <- c(rnorm(30,-3),rnorm(30,3))
x <- cbind(x = tmp,y = rev(tmp))
plot(x)
```
Now we have some made up data in 'x' let's see how kmeans works with this data. 

```{r}
k <-kmeans(x,centers = 2, nstart = 20)
k
```
> Q. How many points are in each cluster? 

```{r}
k$size
```
>Q. How do we get to the cluster membership/assignment? 

```{r}
k$cluster
```

>Q. What about cluster centers?

```{r}
k$centers
```

Now we've got to main results, let's use thme to plot our data with the kmeans result

```{r}
plot(x,col=k$cluster)
points(k$centers,col="blue", pch=15)
```

## Now for Hierarchial Clustering

We will cluster the same data 'x' with the 'hclust()'. In this case 'hclust()' requires a distance matrix as input 

```{r}
hc <-hclust(dist(x))
hc
```

Let's plot our hclus result
```{r}
plot(hc)
```

To get our cluster membershit vector we need to "cut" the tree with 'cutree()' 

```{r}
grps <- cutree(hc,h=8)
grps
```

Now plot our data with hclust() results

```{r}
plot(x,col=grps)
```

# Principal Component Analysis (PCA)

## PCA of UK Food Data

Read data from website and try a few visualization 

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names =1)
x
```



```{r}
cols <-rainbow(nrow(x))
barplot(as.matrix(x),col=cols)
```


```{r}
barplot(as.matrix(x),col=cols, beside = TRUE)
```

```{r}
pairs(x,col=cols)
```

PCA to the rescue!
The main base R PCA  function is called 'prcomp()'and we will need to give it the transporse of our input data. 

```{r}
pca <-prcomp(t(x))
pca
```
```{r}
pca <-prcomp(t(x))
```
There is a nice summary of how well PCA is doing
```{r}
summary(pca)
```

```{r}
attributes (pca)
```

To make our new PCA plot (aka PCA score plot) we access 'pca$x'
```{r}
plot(pca$x[,1],pca$x[,2])
text(pca$x[,1],pca$x[,2],colnames(x))
```
Color up the plot 

```{r}
country_cols <- c("orange", "red", "blue", "green")
plot(pca$x[,1],pca$x[,2],xlab="PC1", ylab="PC2")
text(pca$x[,1],pca$x[,2],colnames(x), col=country_cols)
```

Calculate how much variation in the original data each PC accounts for
```{r}
v <- round( pca$sdev^2/sum(pca$sdev^2) * 100 )
v
```
Or second row 
```{r}
z <- summary(pca)
z$importance
```
Plot of variances(eigenvalues) with respect to the principal component number (eigenvector number)
```{r}
barplot(v, xlab="Principal Component", ylab="Percent Variation")
```
Consider the influence of each original variable upon the principal components (loading scores)
```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,1], las=2 )
```

>Q. Generate a similar 'loadings plot' for PC2. What two food groups feature prominately and what does PC2 mainly tell us about? 


```{r}
par(mar=c(10, 3, 0.35, 0))
barplot( pca$rotation[,2], las=2 )
```

Fresh_potatoes and Soft_drinks are the two prominent groups. PC2 tells us that Soft_drinks push the other countries to the left negative side of the plot while Fresh_potatoes pushes N.Ireland to the right positive side of the plot. 

Inbuilt biplot()

```{r}
biplot(pca)
```



##PCA of RNA-Seq data 

Read in data from website
```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)

```

>Q How many genes are in this data set?

```{r}
pca <-prcomp(t(rna.data))
```
There is a nice summary of how well PCA is doing
```{r}
summary(pca)
```

Do our PCA plot of this RNA-seq data
```{r}
plot(pca$x[,1],pca$x[,2], xlab="PC1", ylab="PC2")
```
```{r}
plot(pca$x[,1],pca$x[,2], xlab="PC1", ylab="PC2")
text(pca$x[,1],pca$x[,2],colnames(rna.data))
```

Barplot summary of Proportion of Variance for each PC 
```{r}
plot(pca, main="Quick scree plot")
```

Make a scree plot ourselves

```{r}
## Variance captured per PC 
pca.var <- pca$sdev^2

## Percent variance is often more informative to look at 
pca.var.per <- round(pca.var/sum(pca.var)*100, 1)
pca.var.per
```
```{r}
barplot(pca.var.per, main="Scree Plot", 
        names.arg = paste0("PC", 1:10),
        xlab="Principal Component", ylab="Percent Variation")
```
Make the PCA plot more attractive 
```{r}
## A vector of colors for wt and ko samples
colvec <- colnames(rna.data)
colvec[grep("wt", colvec)] <- "red"
colvec[grep("ko", colvec)] <- "blue"

plot(pca$x[,1], pca$x[,2], col=colvec, pch=16,
     xlab=paste0("PC1 (", pca.var.per[1], "%)"),
     ylab=paste0("PC2 (", pca.var.per[2], "%)"))

text(pca$x[,1], pca$x[,2], labels = colnames(rna.data), pos=c(rep(4,5), rep(2,5)))
```
Using ggplot 
```{r}
library(ggplot2)

df <- as.data.frame(pca$x)

# Our first basic plot
ggplot(df) + 
  aes(PC1, PC2) + 
  geom_point()
```
Add a condition specific color and labels for aesthetics
```{r}
df$samples <- colnames(rna.data) 
df$condition <- substr(colnames(rna.data),1,2)

p <- ggplot(df) + 
        aes(PC1, PC2, label=samples, col=condition) + 
        geom_label(show.legend = FALSE)
p
```

```{r}
p + labs(title="PCA of RNASeq Data",
       subtitle = "PC1 clealy seperates wild-type from knock-out samples",
       x=paste0("PC1 (", pca.var.per[1], "%)"),
       y=paste0("PC2 (", pca.var.per[2], "%)"),
       caption="BIMM143 example data") +
     theme_bw()
```



