---
title: "class10"
author: "Katie Chau"
date: "2/21/2022"
output:
  pdf_document: default
  html_document: default
---

## Population Analysis

How many samples do we have?
```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```

```{r}
nrow(expr)
```
```{r}
table(expr$geno)
```

>Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes. Hint: The read.table(),summary() and boxplot() functions will likely be useful here. There is an example R script online to be used ONLY if you are struggling in vein. Note that you can find the medium value from saving the output of the boxplot() function to an R object and examining this object. There is also the medium() and summary() function that you can use to check your understanding.


```{r}
library(ggplot2)
```

Let's make a boxplot
```{r}
n_fun <- function(geno){
  return(data.frame(y = median(geno), label = paste0("median = ",length(geno))))
}
ggplot(expr)+aes(x=geno,y=exp,fill=geno)+
  geom_boxplot(notch=TRUE)+
 stat_summary(fun.data = n_fun, geom = "text")

```
I followed the lab walk through, but I'm still not sure how you get the medium() or median values. There is no medium function available so I looked online to find help. I also could not find the example R script. 



>Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3? Hint: An example boxplot is provided overleaf – yours does not need to be as polished as this one.

The A/A expression levels are higher than the G/G expression levels by about 10 on average. Yes, I believe the SNP does effect the expression of ORMDL3.   

